Modelling RNA-based oscillatory circuits

### 2015-09-29 Presentation/MscPresentation.tex
more presentation work
final commits

\begin{frame}{Objectives}{}
  \begin{itemize}
    \item  Rational design of genetic circuits now possible
      \begin{itemize}
      \item ODE models used for prediction.
      \item Model parameters unknown.
      \end{itemize}
    \item Project studied a recently designed genetic circuit, attempted parameter estimation.
    \item Parameter estimates would inform future design/experiments.
        \end{itemize}
\end{frame}

@@

\begin{frame}{ODE model for system (2) - Translation step}
\begin{columns}
\column{.4\textwidth}
\includegraphics[trim = 0 0 0 0,clip = true,scale = 0.31]{Figures/schematic_translation}
\column{.6\textwidth}
  \begin{itemize}
    \item Translation of stable complex \textcolor{blue}{($c$)} into immature GFP \textcolor{blue}{($p$)}. 
    \item maturation of GFP \textcolor{blue}{($g$)}, machine calibration giving measured fluorescence ($z$).
    \end{itemize}
\footnotesize
\begin{align*} 
\frac{d\textcolor{blue}{p}}{dt} & = \beta m +f_{s}\beta c -(\gamma + \mu + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
\frac{d\textcolor{blue}{g}}{dt} & = \gamma p - (\mu + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \\
z &= z_{0} +\frac{g}{\Theta} 
\end{align*}
\end{columns}
\end{frame}

@@

\begin{frame}{Recent experimental data}{}
   \includegraphics[trim = 0 300 0 300,clip = true,scale = 0.55]{Figures/13_9}
  \begin{itemize}
    \item Single cell fluoresence time series data (Jaramillo Lab, unpublished).
    \item  Data records $z= z_{0} +\frac{g}{\alert{\Theta}} $
    \end{itemize}
\end{frame}

@@

\begin{frame}{Initial estimation results}
  \includegraphics[scale = 0.27, clip = true, trim = 130 0 0 50]{Figures/13_9_hist}
  \vspace{-10mm}
        \begin{itemize}
    \item Similar error values for all results.
    \item Many parameters poorly estimated.
            \end{itemize}
\end{frame}

@@

\small
\begin{frame}{Full model, with parameters to be estimated}
\begin{align*} 
\frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\alert{\mu} + \alert{\delta_{s}})s -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m \\
\frac{dm}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\alert{\mu} + \alert{\delta_{m}})m -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m  \\
\frac{ds:m}{dt} & = \alert{k_{\mathrm{on}}}sm  - (\alert{k_{\mathrm{off}}}+ \alert{k_{\mathrm{hyb}}})s:m  -(\alert{\mu} + \alert{\delta_{m}} )s:m \\
\frac{dc}{dt} & = \alert{k_{\mathrm{hyb}}}s:m  -(\alert{\mu} + \alert{\delta_{m}})c  
\end{align*}
\begin{center}
\rule{0.5\textwidth}{.4pt}
\end{center}
\begin{align*}
\frac{dp}{dt} & = \alert{\beta} m +\alert{f_{s}}\beta c -(\gamma + \alert{\mu} + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
\frac{dg}{dt} & = \gamma p - (\alert{\mu} + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \\
z &= z_{0} +\frac{g}{\alert{\Theta}} 
\end{align*}
\end{frame}
\normalsize

@@

\begin{frame}{Simplified model results (1)}
  \includegraphics[scale = 0.29, clip = true, trim = 100 0 0 400]{Figures/13_9_hist_simplified}
  \begin{itemize}
\item  Error values as low as full model
\item Clearer parameter estimation results
\item Two peak structure due to local minima
\end{itemize}
\end{frame}

@@

\begin{frame}{Future work}
  \begin{itemize}
  \item Fluorescence data \alert {not enough} to estimate all unknown parameters 
  \item Further experiments needed - observe fast timescale.
  \item Investigate methods of McAuley et al. 
  \item Change methodology - Bayesian Methods?
  \end{itemize}
  \end{frame}

@@

>>>>

\begin{frame}{Objectives}{}
  \begin{itemize}
  \item Goal in synthetic biology - manipulate natural cellular machinery
    \item  Rational design of genetic circuits now possible
      \begin{itemize}
      \item ODE models used for prediction.
      \item Model parameters unknown.
      \end{itemize}
    \item Project studied a recently designed genetic circuit 
    \item \textcolor{blue}{Aim}: use \textcolor{blue}{ODE model} with \textcolor{blue}{recent data} to perform \textcolor{blue}{parameter estimation}.
    \item Parameter estimates would inform future design/experiments.
        \end{itemize}
\end{frame}

@@

\begin{frame}{ODE model for system (2) - Translation step}
\begin{columns}
\column{.4\textwidth}
\includegraphics[trim = 0 0 0 0,clip = true,scale = 0.31]{Figures/schematic_translation}
\column{.6\textwidth}
  \begin{itemize}
    \item Translation of stable complex \textcolor{blue}{($c$)} into immature GFP \textcolor{blue}{($p$)}. 
    \item maturation of GFP \textcolor{blue}{($g$)}, machine calibration giving measured fluorescence \textcolor{blue}{($z$)}.
    \end{itemize}
\footnotesize
\begin{align*} 
\frac{d\textcolor{blue}{p}}{dt} & = \beta m +f_{s}\beta c -(\gamma + \mu + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
\frac{d\textcolor{blue}{g}}{dt} & = \gamma p - (\mu + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \\
z &= z_{0} +\frac{g}{\Theta} 
\end{align*}
\end{columns}
\end{frame}

@@

\begin{frame}{Recent experimental data}{}
   \includegraphics[trim = 0 300 0 300,clip = true,scale = 0.55]{Figures/13_9}
  \begin{itemize}
      \item  Data records $z= z_{0} +\frac{g}{\alert{\Theta}} $
    \item Single cell fluoresence time series data (Jaramillo Lab, unpublished).
    \end{itemize}
\end{frame}

@@

\begin{frame}{Initial estimation results}
  \includegraphics[scale = 0.27, clip = true, trim = 100 0 0 40]{Figures/13_9_hist}
  \vspace{-10mm}
        \begin{itemize}
            \item Many parameters poorly estimated.
    \item Similar error values for all results.
            \end{itemize}
\end{frame}

@@

\small
\begin{frame}{Plotting all states of model output}
\begin{align*} 
\frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\alert{\mu} + \alert{\delta_{s}})s -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m \\
\frac{dm}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\alert{\mu} + \alert{\delta_{m}})m -\alert{k_{\mathrm{on}}}sm +\alert{k_{\mathrm{off}}}s:m  \\
\frac{ds:m}{dt} & = \alert{k_{\mathrm{on}}}sm  - (\alert{k_{\mathrm{off}}}+ \alert{k_{\mathrm{hyb}}})s:m  -(\alert{\mu} + \alert{\delta_{m}} )s:m \\
\frac{dc}{dt} & = \alert{k_{\mathrm{hyb}}}s:m  -(\alert{\mu} + \alert{\delta_{m}})c  
\end{align*}
\begin{center}
\rule{0.5\textwidth}{.4pt}
\end{center}
\begin{align*}
\frac{dp}{dt} & = \alert{\beta} m +\alert{f_{s}}\beta c -(\gamma + \alert{\mu} + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}   \\
\frac{dg}{dt} & = \gamma p - (\alert{\mu} + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \\
z &= z_{0} +\frac{g}{\alert{\Theta}} 
\end{align*}
\end{frame}
\normalsize

@@

\begin{frame}{Simplified model results (1)}
  \includegraphics[scale = 0.29, clip = true, trim = 70 0 0 400]{Figures/13_9_hist_simplified}
  \begin{itemize}
\item  Error values as low as full model
\item Clearer parameter estimation results
\item Two peak structure due to local minima
\end{itemize}
\end{frame}

@@

\begin{frame}{Conclusions and further work}
  \begin{itemize}
  \item Fluorescence data \alert {not enough} to estimate all unknown parameters 
  \begin{itemize}
  \item Simplified model not helpful in estimating original parameters.
  \end{itemize}
  \end{itemize}
To get these parameters ...
  \begin{itemize}
  \item Further experiments needed - observe fast timescale.
  \item Improve methodology
    \begin{itemize}
    \item Algorithms for dealing with complex chemical models \& limited data.
    \item Bayesian Methods
      \end{itemize}
  \end{itemize}
  \end{frame}

@@

  \begin{frame}{Additional Slides}
  \Huge
	Additional Slides
	\normalsize
  \end{frame}
